---
title: "Inference for a mean"
editor: visual
---

# Warm up

## Announcements

-   No class tomorrow
-   Lab 8 due Thursday
-   Lab 9 optional
-   Friday we will talk about final report formatting

# Inference for one mean

## Today's focus

-   Inference for a population mean

-   Specifically, estimating a population mean via a confidence interval

-   And generally, **what in the world is a confidence interval?!**

    -   Why is it constructed?

    -   How is it constructed?

    -   What do the bounds of a confidence interval mean?

    -   What does the confidence level mean?

    -   What is the margin of error?

## Setup

```{r}
#| label: load-packages
#| message: false
#| 
library(tidyverse)
library(openintro)
library(tidymodels)
```

## Bootstrap confidence interval for the mean {.smaller}

1.  Take a bootstrap sample (sample with replacement) of size n (the original sample size) from the original sample

2.  Record the mean

3.  Repeat steps 1 and 2 many times to build the bootstrap distribution

4.  Find the bootstrap confidence interval using one of two methods:

    -   Percentile method: The bounds of the middle X% of the bootstrap distribution comprise the X% bootstrap interval.

    -   Standard error method:

        -   Calculate the standard error for the bootstrap distribution

        -   Find the critical value associated with the X% confidence interval

        -   Calculate the **margin of error** (ME) as the critical value $\times$ standard error

        -   Construct the confidence interval a the original sample mean $\pm$ ME

## Case study: Length of gestation {.smaller}

::: task
Every year, the United States Department of Health and Human Services releases to the public a large dataset containing information on births recorded in the country. This dataset has been of interest to medical researchers who are studying the relation between habits and practices of expectant mothers and the birth of their children. In this case study we work with a random sample of 1,000 cases from the dataset released in 2014. The length of pregnancy, measured in weeks, is commonly referred to as gestation. The histogram below shows the distribution of lengths of gestation from the random sample of 1,000 births.
:::

```{r}
#| output-location: column

ggplot(
  openintro::births14, 
  aes(x = weeks)
  ) +
  geom_histogram(binwidth = 1) +
  labs(
    x = "Gestation (weeks)",
    y = "Count",
    title = "Random sample of 1,000 births"
  )
```

## CS: Length of gestation - `boot_dist`

::: task
The distribution of bootstrapped means of gestation from 1,500 different bootstrap samples.
:::

```{r}
#| output-location: column

set.seed(1234)

boot_dist <- births14 |>
  specify(response = weeks) |>
  generate(
    reps = 1500, 
    type = "bootstrap"
  ) |>
  calculate(stat = "mean")

visualize(boot_dist)
```

## CS: Length of gestation - percentile interval

::: task
What does the following code do? How do you adjust it to change the confidence level?
:::

```{r}
boot_dist |>
  get_ci()


boot_dist |>
  summarize(
    lower = quantile(stat, 0.025),
    upper = quantile(stat, 0.975)
  )
```

## CS: Length of gestation - standard error interval {.smaller}

::: task
What does the following code do? Why do we need to provide `point_estimate`?
:::

```{r}
obs_stat <- births14 |>
  specify(response = weeks) |>
  calculate(stat = "mean")

obs_stat

boot_dist |>
  get_ci(
    type = "se",
    point_estimate = obs_stat,
    level = 0.95
  )
```

## Interpretations

-   The interval: (38.5, 38.8)

-   The confidence level: 95%

::: task
What does 95% confidence interval mean?
:::

> We are 95% confident that the true mean gestational period is between 38.5 and 38.8 weeks.

# Bootstrap confidence intervals

## Exercise 1

Construct and interpret a 95% confidence interval, using the standard error method with the `get_ci()` function, for the average length of gestation.

```{r}
#| label: se-95
obs_stat <- births14 |>
  specify(response = weeks) |>
  calculate(stat = "mean")

obs_stat

boot_dist |>
  get_ci(
    type = "se",
    point_estimate = obs_stat,
    level = 0.95
  )

```

*Add response here.*

## Exercise 2

Construct the interval **without** using the `get_ci()` (or `get_confidence_interval()`) function.

```{r}
#| label: se-95-again
# option 1 for se
se <- boot_dist |>
  summarize(se = sd(stat)) |>
  pull(se)

se

# option 2 for se
sd(boot_dist$stat)


#option 1 for ci

obs_stat |>
  mutate(lower = stat - 1.96*se,
         upper = stat + 1.96*se)

# option 2 for ci

xbar = mean(births14$weeks)

c(xbar - 1.96*se, xbar + 1.96*se)
```

## Exercise 3

Would you expect a 90% confidence interval for the average length of gestation to be wider or narrower? Explain your reasoning.

## Exercise 4

Now construct a 90% confidence interval for the average length of gestation and confirm your answer from the previous exercise. Repeat as little of your code as possible.

```{r}
#| label: se-90


boot_dist |>
  get_ci(
    type = "se",
    point_estimate = obs_stat,
    level = 0.90
  )
```

# Confidence intervals with mathematical models

If technical conditions are satisfied, the distribution of the sampling mean (i.e., the sampling distribution) follows the t-distribution with n - 1 degrees of freedom, where n is the sample size.

```{r}
sampling_dist <- births14 |>
  specify(response = weeks) |>
  assume(distribution = "t")

sampling_dist

visualize(sampling_dist)
```

## Exercise 5

Construct and visualize a confidence interval using mathematical models for the average length of gestation.

```{r}
#| label: math-95

# option

 births14 |>
   specify(response = weeks) |>
   assume(distribution = "t") |>
   get_ci(type = "se", point_estimate =  obs_stat)
 
 
 #option 2
 s <- sd(births14$weeks)
 n <- nrow(births14)
 se <- s/sqrt(n)
 se
 
 # mathematical model ci approx N 
 

# cv using normal
qnorm(0.975)

# cv using t dist
qt(0.975, df = n-1)

#our case: df is large, so both are about 1.96

c(xbar - 1.96*se, xbar + 1.96*se)

```

