---
title: "Regression with multiple predictors "
editor: visual
---

# Warm up

## Announcements {.smaller}

-   No lab or class on Monday, May 27th.
-   Project proposal due Tuesday, May 28th.
-   Lab 3 due Wednesday, May 29th.



## Load packages and data

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(tidymodels)
library(scatterplot3d)
```

# $R^2$ and model fit

## $R^2$

$R^2$, aka "the coefficient of determination" or "correlation squared" is a way to see how well a given model fits the data.

$$
R^2 = r_i^2 
$$

## Sum of squares total, SST

The sum of squares total is a measure of the total variability in the outcome variable:

$$
SST = (y_1 - \bar{y})^2 + (y_2 - \bar{y})^2 + \cdots + (y_n - \bar{y})^2
$$

## Sum of squares residuals, SSE

The sum of squares residuals (error) is a measure of the variability in the residuals, i.e., variability left unexplained in the outcome variable after the model is fit:

$$
SSE = (y_1 - \hat{y}_1)^2 + (y_2 - \hat{y}_2)^2 + \cdots + (y_n - \hat{y}_n)^2
$$

## $R^2$, another look

$$
R^2 = \frac{SST - SSE}{SST} = 1 - \frac{SSE}{SST}
$$

. . .

-   This can be summarized as "$R^2$ is 1 minus the sum of squared residuals divided by the sum of squared total".

. . .

-   In other words, $R^2$ is the proportion of variability in the outcome that is explained by the model.

## $R^2$

::: incremental
-   If the sum of squared residuals is 0, then the model explains all variability and $R^2 = 1 - 0 = 1$.

-   If the sum of squared residuals is the same as all the variability in the data, then model does not explain any variability and $R^2 = 1 - 1 = 0$.

-   $R^2$ is a measure of the proportion of variability the model explains. An $R^2$ of 0 is a poor fit and $R^2$ of 1 is a perfect fit.
:::

## Finding $R^2$

To find $R^2$ simply call the function `glance()` on your `model_fit`, e.g.

```{r}
#| eval: false

model_fit <- linear_reg() |>
  fit(outcome ~ predictor, data = data_set)
  
glance(model_fit)
```

# Models with multiple predictors

## Two predictors {.smaller}

$$
y = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \epsilon
$$

::: incremental
-   $y$: the **outcome** variable. Also called the "response" or "dependent variable". In prediction problems, this is what we are interested in predicting.
-   $x_i$: the $i^{th}$ **predictor**. Also commonly referred to as "regressor", "independent variable", "covariate", "feature", "the data".
-   $\beta_i$: "constants" or **coefficients** i.e. fixed numbers. These are **population parameters**. $\beta_0$ has another special name, "the intercept".
-   $\epsilon$: the **error**. This quantity represents observational error, i.e. the difference between our observation and the true population-level expected value: $\beta_0 + \beta_1 x$.
-   Effectively this model says our data $y$ is linearly related to the $x_1$ and $x_2$ but is not perfectly observed due to unexplained errors.
:::

## A simple example

Let's examine the first quarter of 2020 high prices of Microsoft, IBM, and Apple stocks to illustrate some ideas.

```{r}
#| label: simple-example
#| echo: false
#| warning: false

stocks <- read_csv("stocks.csv")

stocks_subset <- stocks |>
  slice(1:62)

scatterplot3d(
  stocks_subset[,c("MSFT.High", "IBM.High", "AAPL.High")],
  pch = 19, 
  color="steelblue",
  xlab = "MSFT High $", 
  ylab = "IBM High $", 
  zlab = "AAPL High $"
) 
```

-   If we have three measurements (variables) then each observation is a point in three-dimensional space. In this example, we can choose one of our measurements to be the outcome variable (e.g. Apple stock price) and use our other two measurements (MSFT and IBM price) as predictors.

-   In general, the total number of measurements, i.e. variables (columns) in our linear model represents the spatial dimension of our model.

## 2 predictors + 1 outcome = 3 dimensions

The fitted linear model no longer looks like a line, but instead looks like a **plane**. It shows our prediction of AAPL price ($y$) given both MSFT price ($x_1$) and IBM price ($x_2$).

```{r}
#| label: simple-example-plane
#| echo: false
#| warning: false

plot3stocks <- scatterplot3d(
  stocks_subset[,c("MSFT.High", "IBM.High", "AAPL.High")],
  pch = 19, 
  color="steelblue",
  xlab = "MSFT High $", 
  ylab = "IBM High $", 
  zlab = "AAPL High $"
)

model <- lm(AAPL.High ~ MSFT.High + IBM.High, data = stocks_subset)
plot3stocks$plane3d(model, col = "darkred")
```

## Fitting a multiple regression model in R

Find the **equation of the plane** by adding in new predictors:

```{r}
#| eval: false

my_model_fit <- linear_reg() |>
  fit(outcome ~ predictor1 + predictor2 + predictor3 + ..., data = data_frame)
```

. . .

-   This code template will fit the model according to the ordinary least squares (OLS) objective function, i.e., we are finding the equation of the hyperplane that minimizes the sum of squared residuals

. . .

-   You can then display a tidy output of the model with the `tidy()` function on your fitted model: `tidy(my_model_fit)`


Today we'll explore the question "How do volume and weights books relate?" and "How, if at all, does that change when we take whether the book is hardback or paperback into consideration?"

# Goals

-   Build, fit, and interpret linear models with more than one predictor

-   Use categorical variables as a predictor in a model

-   Compute $R^2$ and use it to select between models

# Packages

```{r}
#| label: load-packages2
#| message: false

library(DAAG)
```

# Data

The data for this application exercise comes from the `allbacks` dataset in the **DAAG** package. The dataset has `r nrow(allbacks)` observations and `r ncol(allbacks)` columns. Each observation represents a book. Let's take a peek at the data:

```{r}
#| label: peek-data

allbacks
```

Note that `volume` is measured in cubic centimeters and `weight` is measured in grams. More information on the dataset can be found in the documentation for `allbacks`, with `?allbacks`.

# Single predictor

## Exercise 1

Visualize the relationship between volume (on the x-axis) and weight (on the y-axis). Overlay the line of best fit. Describe the relationship between these variables.

```{r}
#| label: plot-weight-volume

# add code here
```

> Add response here.

## Exercise 2

Fit a model predicting weight from volume for these books and save it as `weight_fit`. Display a tidy output of the model.

```{r}
#| label: fit-weight-volume

# add code here
```

## Exercise 3

Interpret the slope and the intercept in context of the data.

> Add response here.

## Exercise 4

Calculate the $R^2$ of this model and interpret it in context of the data.

```{r}
#| label: glance-weight-volume

# add code here
```

> Add response here.

# Multiple predictors

## Exercise 5

Visualize the relationship between volume (on the x-axis) and weight (on the y-axis), taking into consideration the cover type of the book. Use different colors and shapes for hardback and paperback books. Also use different colors for lines of best fit for the two types of books. In addition, add the overall line of best fit (from Exercise 1) as a gray dashed line so that you can see the difference between the lines when considering and not considering cover type.

```{r}
#| label: plot-weight-volume-cover

# add code here
```

## Exercise 6

Fit a model predicting weight from volume for these books and save it as `weight_cover_fit`. Display a tidy output of the model.

```{r}
#| label: fit-weight-volume-cover

# add code here
```

## Exercise 7

In the model output we have a variable `coverpb`. Why is only the `pb` (paperback) level of the `cover` variable shown? What happened to the `hb` (hardback) level?

> Add response here.

## Exercise 8

Interpret the slopes and the intercept in context of the data.

> Add response here.

## Exercise 9

First, guess whether the $R^2$ of this model will be greater than, less than, or the same as the previous model, or whether we can't tell. Then, calculate the $R^2$ of this model to confirm your guess, and then interpret it in context of the data.

```{r}
#| label: glance-weight-volume-cover

# add code here
```

> Add response here.

## Exercise 10

Which model is preferred for predicting book weights and why?

> Add response here.

## Exercise 11

Using the model you chose, predict the weight of a hardcover book that is 1000 cubic centimeters (that is, roughly 25 centimeters in length, 20 centimeters in width, and 2 centimeters in height/thickness).

```{r}
#| label: predict

# add code here
```

