---
title: "Mathematical models for proportions"
editor: visual
---

# Warm up

## Announcements

-   No classes on Wednesday
-   Lab 8 due Thursday
-   Lab 9 optional
-   Friday we will talk about final report formatting

# Packages and data

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(openintro)
```

# Single proportion hypothesis testing

According to [Duke Facts](https://facts.duke.edu/), 16% of Duke students are North Carolina residents.Suppose we take a random sample of 500 Duke students and find that 100 are NC residents. Can we conclude that the proportion is greater than 16%?

::: task
How would you perform a hypothesis test to test whether the true proportion of Duke students that are NC residents is truly smaller than 16%?
:::

## Exercise 1

What are the hypotheses?

## No dataset

We have explored how to use a data set to simulate sampling distribution under null and perform randomization test.

```{r}
#| label: randomization-test-if-data
#| eval: false

set.seed(1234)

null_dist <- duke_nc |>
  specify(response = resident, success = "NC") |>
  hypothesise(null = "point", p = 0.12) |>
  generate(reps = 1000, type = "draw") |>
  calculate(stat = "prop")


null_dist |>
  get_p_value(obs_stat, direction = "greater")


visualize(null_dist) +
  shade_p_value(obs_stat, direction = "greater")

```

We don't have a data set to run this code!

## Exercise 2

What is the observed statistic?

```{r}
#| label: obs_stat


```

## Null distribution

We want to simulate null distribution, i.e., we want to randomly sample 500 observations (students) from a population (all Duke students) where the probability of being a NC resident is 0.16.

The **binomial distribution** is a special discrete distribution where there are two distinct complementary outcomes, a "success" and a "failure" ("NC" and "not NC" in our example). Binomial distributions takes in two parameters: $n$ - number of trials/sample size, and $p$ - probability of success. The $n$ trials/observations are independent. That is, the outcome of any trial does not affect the outcome of the others. Think of it as flipping a coin that has probability $p$ of landing on heads $n$ times and recording the number of heads observed.

::: task
In our example, what is $n$ and what is $p$?
:::

## Simulate null distribution

We will draw samples of size 500 each observation having a 0.16 probability of being from NC and will record the proportion that came from NC. We will do this 1000 times.

```{r}
#| label: null-dist

set.seed(1234)

nc_count <- tibble(stat = rbinom(n = 1000, size = 500, p = 0.16))
null_dist_nc = tibble(stat = nc_count$stat/500)

#show first 5 results
nc_count[1:5,]
null_dist_nc[1:5,]
```

We can visualize null distribution:

```{r}
#| label: null-dist-plot
#| eval: false

ggplot(null_dist_nc, aes(x = stat)) +
  geom_histogram(bins = 15, color = "white") + 
  geom_vline(xintercept = 0.2, color = "red" )

```

## Calculate p-value

How do we calculate p-value?

```{r}
#| label: p-value

```

## Exercise 3

Use the simulated data to perform randomization test using count instead of proportion. Define your new hypotheses, observed statistic, visualize null distribution, and calculate the p-value.

```{r}
#| label: count-test
```

::: task
Why are we using proportion and not count for null distribution?
:::

# Mathematical models

In CLT lecture, we introduced the normal distribution and showed how it can be used as a mathematical model to describe the variability of a statistic.There are conditions under which a sample proportion $\hat p$ is well modeled with a normal distribution.

::: callout-important
The sampling distribution for $\hat p$ (the sample proportion) based on a sample of size $n$ from a population with true proportion $p$ is nearly normal when:

1.  The sample's observations are independent, e.g., are from a simple random sample.
2.  We expected to see at least 10 successes and 10 failures in the sample, i.e., $np \geq 10$ and $n(1-p)\geq 10$. This is called the **success-failure condition**.

When these conditions are met, them the sampling distribution of $\hat p$ is nearly normal with mean $p$ and standard error of $\hat p$ as $SE = \sqrt{\frac{p(1-p)}{n}}$.
:::

Typically we do not know the true proportion $p$ so we substitute some value to check conditions and estimate the standard error. For confidence intervals, the sample proportion $\hat p$ is used to check the success-failure condition and compute the standard error. For hypothesis tests, typically the null value – that is, the proportion claimed in the null hypothesis – is used in place of $p$.

## Exercise 4

Is the success-failure condition satisfied in our example of Duke students from NC?

# Confidence interval for a proportion

When $\hat p$ can be modeled using a normal distribution, the confidence interval for $p$ takes the form $\hat p \pm z^* \times SE$. The distance given by $z^* \times SE$ is called the margin of error. Since $p$ is unknown, we use $\hat p$ to approximate the standard error: $SE \approx \sqrt{\frac{\hat p(1-\hat p)}{n}}$.

## $z^*$

The value $z^*$ determines the confidence level. In particular, if you are interested in $(1-\alpha)\times 100$% confidence interval, $z^*$ describes how many standard deviations away the bounds of $(1-\alpha)\times 100$% on standard normal (mean 0 and standard deviation 1) are. For example, common value of $z^*$ is 1.96 (often approximated to be 2) indicating that the margin of error describes the variability associated with 95% of the sampled statistics.

The value of $z^*$ can be found using `qnorm()` function in R.

```{r}
#| label: z_stars

# 99.7% CI
qnorm(0.0015)
# 95% CI
qnorm(0.025)
# 68% CI:
qnorm(0.16)

```

::: task
How do these relate to 68-95-99.7 rule?
:::

## Exercise 5

Construct a 95% confidence interval for $p$, the proportion of Duke students who are residents of North Carolina.

# Hypothesis test for proportion

Suppose we have $H_0: p = p_0$ vs $H_A: p\neq p_0$ where $p_0$ is the null value we are testing (0.16 in our example).

To apply the normal distribution framework in the context of a hypothesis test for a proportion, the independence and success-failure conditions must be satisfied. In a hypothesis test, the success-failure condition is checked using the null proportion: we verify $np_0$ and $n(1-p_0)$ are at least 10, where $p_0$ is the null value.

Once conditions for using normal distribution for sample proportion under $H_0$ are verified, we calculate standard error using $p_0$. Thus, we assume, under null, sample proportions is distributed normally with mean $p_0$ and standard error $\sqrt{\frac{p_0(1-p_0)}{n}}$ (if $H_0$ is true, this is just CLT).

## Exercise 6

Perform the hypothesis test testing whether the true proportion of Duke students that are NC residents is truly smaller than 16%?

```{r}
#| label: math-test
#| eval: false
null_mean = 
null_se = 
pnorm(q = obs_stat,mean = null_mean, sd = null_se )
```
