---
title: "Donors and yawners"
categories: 
  - Application exercise
editor: visual
editor_options: 
  chunk_output_type: console
---

In this application exercise, we'll introduce pipelines for conducting hypothesis tests with randomization.

# Goals

-   Conduct a hypothesis test for a proportion

-   Conduct a hypothesis test for a difference in proportions

# Packages and data

We'll use the **tidyverse** and **tidymodels** packages as usual and the **openintro** package for the datasets.

```{r}
#| label: load-packages-data
#| message: false

library(tidyverse)
library(tidymodels)
library(openintro)
```

# Case study 1: Donors

The first dataset we'll use is `organ_donors`, which is in your `data` folder:

```{r}
organ_donor <- read_csv("data/organ-donor.csv")
```

The hypotheses we are testing are:

$H_0: p = 0.10$

$H_A: p \ne 0.10$

where $p$ is the true complication rate for this consultant.

## Exercise 1

Construct the null distribution with 100 resamples. Name it `null_dist_donor`. How many rows does `null_dist_donor` have? How many columns? What does each row and each column represent?

```{r}
#| label: null-dist-donor

# add code here
```

*Add response here.*

## Exercise 2

Where do you expect the center of the null distribution to be? Visualize it to confirm.

```{r}
#| label: null-dist-donor-plot

# option 1

# add code here

# option 2

# add code here
```

## Exercise 2

Calculate the observed complication rate of this consultant. Name it `obs_stat_donor`.

```{r}
#| label: obs-stat-donor

# add code here
```

## Exercise 3

Overlay the observed statistic on the null distribution and comment on whether an observed outcome as extreme as the observed statistic, or lower, is a likely or unlikely outcome, if in fact the null hypothesis is true.

```{r}
#| label: obs-stat-donor-plot

# option 1

# add code here

# option 2

# add code here
```

## Exercise 4

Calculate the p-value and comment on whether it provides convincing evidence that this consultant incurs a lower complication rate than 10% (overall US complication rate).

*Add response here.*

```{r}
#| label: pvalue-donor

# option 1

# add code here

# option 2

# add code here
```

## Exercise 5

Let's get real! Redo the test with 15,000 simulations. Note: This can take some time to run.

```{r}
#| label: donor-ht-more-sims

# add code here
```

# Case study 2: Yawners

## Exercise 6

Using the `yawn` dataset in the **openintro** package, conduct a hypothesis test for evaluating whether yawning is contagious. First, set the hypotheses. Then, conduct a randomization test using 1000 simulations. Visualize and calculate the p-value and use it to make a conclusion about the statistical discernability of the difference in proportions of yawners in the two groups. Then, comment on whether you "buy" this conclusion.

*Add response here.*

```{r}
#| label: yawn-ht

# add code here
```
