---
title: "Bone density"
categories: 
  - Application exercise
  - Suggested answers
fig-width: 6
fig-asp: 0.618
---

In this application exercise, we'll introduce and work with the normal distribution.

# Goals

-   Calculate probabilities under the normal curve

-   Understand Z scores

-   Build intuition for hypothesis testing and confidence interval building with the normal distribution

# Packages and data

We'll use the **tidyverse** and **openintro** packages.

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(openintro)
```

# Bone density - population

Suppose the bone density for 65-year-old women is normally distributed with mean 809 $mg/cm^3$ and standard deviation of 140 $mg/cm^3$.

Let $x$ be the bone density of 65-year-old women. We can write this distribution of $x$ in mathematical notation as

$$
x \sim N(\mu = 809, \sigma = 140)
$$

Let's set the parameters of this distribution as objects you can use later.

```{r}
#| label: population-parameters

bone_density_mean <- 809
bone_density_sd <- 140
```

## Exercise 1

Visualize the distribution of bone density of 65-year-old women.

```{r}
#| label: viz-pop

# add code here
```

## Exercise 2

Before typing any code, based on what you know about the normal distribution, what do you expect the median bone density to be?

*Add response here.*

## Exercise 3

What bone densities correspond to Q1 (25th percentile), Q2 (50th percentile), and Q3 (the 75th percentile) of this distribution? Use the `qnorm()` function to calculate these values.

```{r}
#| label: quartiles

# add code here
```

## Exercise 4

The densities of three woods are below:

-   Plywood: 540 $mg/cm^3$

-   Pine: 600 $mg/cm^3$

-   Mahogany: 710 $mg/cm^3$

Let's set these as variables we can use later:

```{r}
#| label: trees

plywood <- 540
pine <- 600
mahogany <- 710
```

What is the probability that a randomly selected 65-year-old woman has bones less dense than Pine?

```{r}
#| label: pine-pop

# add code here
```

## Exercise 5

Would you be surprised if a randomly selected 65-year-old woman had bone density less than Mahogany? What if she had bone density less than Plywood? Use the respective probabilities to support your response.

*Add response here.*

```{r}
#| label: mahogany-plywood-pop

# add code here
```

# Bone density - sampling

Suppose you want to analyze the mean bone density for a group of 10 randomly selected 65-year-old women.

## Exercise 6

Are the conditions for the Central Limit Theorem met?

*Add response here.*

## Exercise 7

What is the shape, center, and spread of the distribution of the mean bone density for a group of 10 randomly selected 65-year-old women?

*Add response here.*

## Exercise 8

What is the probability that the mean bone density for the group of 10 randomly-selected 65-year-old women is less dense than Pine?

```{r}
#| label: pine-sampling

# add code here
```

## Exercise 9

Would you be surprised if a group of 10 randomly-selected 65-year old women had a mean bone density less than Mahogany? What the group had a mean bone density less than Plywood? Use the respective probabilities to support your response.

```{r}
#| label: mahogany-plywood-sampling

# add code here
```

## Exercise 10

Explain how your answers differ in similar sounding earlier vs. later exercises.

*Add response here.*

```{r}
#| label: viz-comparison

# add code here
```
