---
title: "Climate change"
categories: 
  - Application exercise
fig-width: 6
fig-asp: 0.618
editor_options: 
  chunk_output_type: console
---

In this application exercise, we'll do inference for a comparing two paired means and many means.

# Packages

We'll use the **tidyverse**, **tidymodels**, **openintro**, and **ggthemes** packages.

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(tidymodels)
library(openintro)
```

# Paired means

Throughout this application exercise, we'll make use of the `climate70` dataset supplied by the **openintro** package.

```{r}
#| label: glimpse

glimpse(climate70)
```

The question we want to answer is: Do these data provide convincing evidence that the number of days above 90 degrees between 2018 and 1948?

First, let's set the hypotheses.

*Add response here.*

To carry out inference on paired data, we pre-compute the difference between paired values at the beginning of the analysis, and use those differences as our values of interest.

```{r}
#| label: calculate-diff

climate70 <- climate70 |>
  mutate(dx90_diff = dx90_2018 - dx90_1948)

climate70 |>
   select(station, latitude, longitude, contains("dx90"))
```

And then visualize these differences:

```{r}
#| label: plot-diff
ggplot(climate70, aes(x = dx90_diff)) +
  geom_histogram()
```

We calculate the observed statistic in the paired setting in the same way that we would outside of the paired setting. Using `specify()` and `calculate()`:

```{r}
#| label: obs-stat

obs_stat <- climate70 |> 
  specify(response = dx90_diff) |> 
  calculate(stat = "mean")

obs_stat
```

Now, we want to compare this statistic to a null distribution, generated under the assumption that the true difference was actually zero, to get a sense of how likely it would be for us to see this observed difference or something more extreme.

Tests for paired data are carried out via the `null = "paired independence"` argument to `hypothesize()`.

```{r}
#| label: null-dist

set.seed(1129)

null_dist <- climate70 |> 
   specify(response = dx90_diff) |> 
   hypothesize(null = "paired independence") |>
   generate(reps = 1000, type = "permute") |>
   calculate(stat = "mean")
   
visualize(null_dist)
```

For each replicate, `generate()` carries out `type = "permute"` with `null = "paired independence"` by:

-   Randomly sampling a vector of signs (i.e., -1 or 1), probability 0.5 for either, with length equal to the input data, and
-   Multiplying the response variable by the vector of signs, "flipping" the observed values for a random subset of value in each replicate

To get a sense for what this distribution looks like, and where our observed statistic falls, we can use `visualize()`:

```{r}
#| label: p-value-viz

visualize(null_dist) + 
  shade_p_value(obs_stat, direction = "two-sided")
```

Then, we calculate the p-value:

```{r}
#| label: p-value

p_value <- null_dist |>
  get_p_value(obs_stat, direction = "two-sided")

p_value
```
