---
title: "Length of gestation"
categories: 
  - Application exercise
fig-width: 6
fig-asp: 0.618
editor_options: 
  chunk_output_type: console
---

In this application exercise, we'll do inference for a single mean, using simulation-based and mathematical models.

# Packages

We'll use the **tidyverse**, **tidymodels**, and **openintro** packages.

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(tidymodels)
library(openintro)
```

# Data

Every year, the United States Department of Health and Human Services releases to the public a large dataset containing information on births recorded in the country. This dataset has been of interest to medical researchers who are studying the relation between habits and practices of expectant mothers and the birth of their children. In this case study we work with a random sample of 1,000 cases from the dataset released in 2014. The length of pregnancy, measured in weeks, is commonly referred to as gestation.

```{r}
glimpse(births14)
```

# Bootstrap confidence intervals

## Exercise 1

Construct and interpret a 95% confidence interval, using the standard error method with the `get_ci()` function, for the average length of gestation.

```{r}
#| label: se-95

# add code here
```

*Add response here.*

## Exercise 2

Construct the interval **without** using the `get_ci()` (or `get_confidence_interval()`) function.

```{r}
#| label: se-95-again

# add code here
```

## Exercise 3

Would you expect a 90% confidence interval for the average length of gestation to be wider or narrower? Explain your reasoning.

*Add response here.*

## Exercise 4

Now construct a 90% confidence interval for the average length of gestation and confirm your answer from the previous exercise. Repeat as little of your code as possible.

```{r}
#| label: se-90

# add code here
```

## Confidence intervals with mathematical models

If technical conditions are satisfied, the distribution of the sampling mean (i.e., the sampling distribution) follows the t-distribution with n - 1 degrees of freedom, where n is the sample size.

```{r}
sampling_dist <- births14 |>
  specify(response = weeks) |>
  assume(distribution = "t")

sampling_dist

visualize(sampling_dist)
```

## Exercise 5

Construct and visualize a confidence interval using mathematical models for the average length of gestation.

```{r}
#| label: math-95

# add code here
```
