---
title: "Issues candidates should talk more about"
categories: 
  - Application exercise
  - Suggested answers
fig-width: 6
fig-asp: 0.618
editor_options: 
  chunk_output_type: console
---

In this application exercise, we'll do inference tables – categorical variables with many levels.

# Packages

We'll use the **tidyverse** and **tidymodels** packages.

```{r}
#| label: load-packages
#| message: false

library(tidyverse)
library(tidymodels)
library(ggthemes)
```

# Data

A September 16-19, 2023, asked North Carolina voters, among other issues, about issues of equality and women's progress. Specifically, one of the questions asked:

> If you had to choose just one issue that you would like candidates to talk more about in the 2024 campaigns, what would that issue be?
>
> -   Economy
>
> -   Abortion/Reproductive rights
>
> -   Immigration
>
> -   Crime
>
> -   Gun rights/restrictions
>
> -   Something else
>
> -   Don't know

The survey also asked respondents' party affiliation:

> What political party do you most identify with?
>
> -   Democrat
>
> -   Republican
>
> -   Unaffiliated
>
> -   Other

The results of this survey are summarized [in this report](https://www.meredith.edu/meredith-poll/n-c-voter-opinions-on-abortion-law-women-leaders-and-other-current-issues/) and the data can be found in your `data` folder: `candidate-talk.csv`.

# Hypotheses

## Exercise 1

State the hypotheses for evaluating whether the issue of choice is independent of party affiliation.

*H0: Issue of choice and party affiliation are independent.*

*HA: Issue of choice and party affiliation are independent.*

# Data

## Exercise 2

Load the data.

```{r}
#| label: load-data
#| message: false

candidate_talk <- read_csv("data/candidate-talk.csv")
```

## Exercise 3

Create a two-way table of the issues across parties and visualize the frequency distribution.

```{r}
#| label: freq-table

candidate_talk <- candidate_talk |>
  mutate(
    party = fct_relevel(party, "Democrat", "Republican", "Unaffiliated", "Other"),
    issue = fct_relevel(issue, "Abortion/Reproductive rights", "Crime", "Economy", "Gun rights/restrictions", "Immigration", "Something else", "Don't know")
  )

candidate_talk_table <- candidate_talk |>
  count(party, issue) |>
  pivot_wider(names_from = "issue", values_from = "n")

candidate_talk_table
```

```{r}
#| label: freq-plot
#| fig-width: 10
#| fig-asp: 0.4

ggplot(candidate_talk, aes(x = party, fill = issue)) +
  geom_bar() +
  scale_fill_colorblind()
```

## Exercise 4

Which do you think should be the explanatory variable and which the response variable? Accordingly, create a visualization that shows the correct conditional probabilities.

```{r}
#| label: prop-plot
#| fig-width: 10
#| fig-asp: 0.4

ggplot(candidate_talk, aes(x = party, fill = issue)) +
  geom_bar(position = "fill") +
  scale_fill_colorblind()
```

# Testing

## Exercise 5

Calculate the observed sample statistic.

```{r}
#| label: obs-stat

obs_stat <- candidate_talk |>
  specify(response = issue, explanatory = party) |>
  hypothesize(null = "independence") |>
  calculate(stat = "Chisq")

obs_stat
```

## Exercise 6

Conduct the hypothesis test using randomization and visualize and report the p-value.

```{r}
#| label: hyp-test

set.seed(1234)

null_dist <- candidate_talk |>
  specify(response = issue, explanatory = party) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  calculate(stat = "Chisq")

null_dist |>
  get_p_value(obs_stat = obs_stat, direction = "greater")

null_dist |>
  visualize() +
  shade_p_value(obs_stat = obs_stat, direction = "greater")
```

## Exercise 7

What is the conclusion of the hypothesis test?

*With a p-value of approximately 0, which is smaller than the discernability level of 0.05, we reject the null hypothesis. The data provide convincing evidence that there is a relationship between party affiliation and issues voters want candidates to discuss.*
