---
title: "Exam 2 Review"
subtitle: "Suggested answers"
execute:
  echo: false
  warning: false
editor_options: 
  chunk_output_type: console
---

```{r}
library(tidyverse)
library(tidymodels)
```

1.  explanatory: income bracket; response: preferred alcoholic beverage

2.  III and IV

3.  $H_0$: Income bracket and preference for alcoholic beverage are independent.

    $H_A$: Income bracket and preference for alcoholic beverage are dependent.

4.  $E = 250 \times \frac{500}{1000} = 125$

5.  The shaded area below represents the p-value.

```{r}
#| fig-width: 6
#| fig-asp: 0.618

income_alcohol <- read_rds(here::here("exam-review", "data/income-alcohol.rds"))

obs_stat <- income_alcohol |>
  specify(response = alcohol, explanatory = income) |>
  calculate(stat = "Chisq")

set.seed(1234)

null_dist <- income_alcohol |>
  specify(response = alcohol, explanatory = income) |>
  hypothesize(null = "independence") |>
  generate(reps = 1000, type = "permute") |>
  calculate(stat = "Chisq")

null_dist_plot <- ggplot(null_dist, aes(x = stat)) +
  geom_histogram(binwidth = 1, color = "white") +
  geom_histogram(
    data = null_dist |> filter(stat > 4.74-0.22),
    aes(x = stat),
    binwidth = 1, color = "white", fill = "pink", alpha = 0.90
  ) +
  geom_vline(
    xintercept = obs_stat |> pull(stat),
    color = "red", linewidth = 1.5
  ) +
  labs(
    title = "Simulation-Based Null Distribution",
    x = "stat",
    y = "count"
  ) +
  scale_y_continuous(
    limits = c(0, 200), 
    breaks = seq(0, 200, 20), 
    minor_breaks = seq(0, 200, 10)
  )

null_dist_plot
```

6.  Add up the heights of the bars that contain simulated sample statistics with a value of 4.74 or higher and divide by the number of simulations. Approximately `{r} 113+76+51+36+30+19+8+6+7+2+3+0+1` (113+76+51+36+30+19+8+6+7+2+3+0+1) simulations out of 1,000 meet this criteria, resulting in a p-value of approximately `{r} (113+76+51+36+30+19+8+6+7+2+3+0+1)/1000`.

```{r}
#| fig-width: 6
#| fig-asp: 0.618

null_dist_plot +
  stat_bin(aes(x = stat, y = after_stat(count)+10, label = after_stat(count)), geom = "text", binwidth = 1)
```

7.  The p-value is much larger than 0.05, therefore the data do not provide convincing evidence that income bracket and preference for alcoholic beverage are dependent.

8.  Take a random sample of size 25, with replacement, from the original sample. Calculate the proportion of students in this simulated sample who work 5 or more hours. Repeat this process 1000 times to build the bootstrap distribution. Take the middle 95% of this distribution to construct a 95% confidence interval for the true proportion of statistics majors who work 5 or more hours.

9.  (40%, 90%)

10. ..

11. ..

12. 

    (c) We are 95% confident that the mean number of texts per month of all American teens is between 1450 and 1550.

13. 

    (a) I \> II

14. a. Power of the test is the probability of rejecting the null hypothesis when the null hypothesis is false.

    b\. Type 1 error rate is the probability of rejecting the null hypothesis when the null hypothesis is true.

    c\. Type 1 error rate is the probability of failing to reject the null hypothesis when the null hypothesis is false.

    d\. 0.05.

15. a\. False. It will be centered at $p$, the true population proportion.

    b\. False. Increasing the number of bootstrap samples does not affect the variability of the sample statistic.

    c\. True.
