---
title: Logistic regression
subtitle: Lecture 10
format: revealjs
auto-stretch: false
editor_options: 
  chunk_output_type: console
---

# Warm up

## Project

::: task
Any questions about projects?
:::

# From last time

## Application exercise

Go to Posit Cloud and continue the project titled `ae-09-Spam`.

# Logistic regression

## The model

Logit form:

$$
log(\frac{p}{1 - p}) = \beta_0 + \beta_1 X_1 + \beta_2 X_2 + ...
$$

Probability form:

$$
p = \frac{e^{\beta_0 + \beta_1 X_1 + \beta_2 X_2 + ...}}{1 + e^{\beta_0 + \beta_1 X_1 + \beta_2 X_2 + ...}}
$$

## Data + packages {.smaller}

```{r}
#| label: load-data-packages
#| message: false

library(tidyverse)
library(tidymodels)

hp_spam <- read_csv("data/hp-spam.csv") |>
  mutate(type = as.factor(type))

hp_spam |>
  select(type, you, capitalTotal)
```

## Fit model and display output

```{r}
spam_fit <- logistic_reg() |>
  fit(type ~ you + capitalTotal, data = hp_spam)
  
tidy(spam_fit)
```

## The model

```{r}
#| echo: false

tidy(spam_fit)
```

$$
log(\frac{p}{1 - p}) = -1.50 + 0.361 \times you + 0.00173 \times  capitalTotal + \epsilon
$$

$$
log(\frac{\hat{p}}{1 - \hat{p}}) = -1.50 + 0.361 \times you + 0.00173 \times  capitalTotal
$$

where $p$ is the probability

# Exam 1

## Exam 1

-   Tally your points on your in class exams and confirm my totals or let me know **now** if anything is off

-   Regrade request policy: <https://sta101-f23.github.io/course-syllabus.html#regrade-requests>

-   Modeling question $\rightarrow$ extra credit question on Exam 2 addressing the same learning goals

-   Statistics!
